---
title: "Compare StepOne vs QuantaStudio for Ssid samples"
author: "AnaPalacio"
date: "Jan 26, 2018"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig_width= 5, fig.height =3)
```

General project set-up 

```{r libraries , results="hide"}
# Get all libraries and sources required to run the script
        source("STEPoneFunction.R")# R.Cunning steponeR function
        library(plyr)
        library(dplyr)
        library(reshape2)
        library(ggplot2)
        library(lme4)
        library(ggthemes)
      
```

# 1. Calculate qPCR RATIOS (Symbiont/Coral) 

Get the raw data for Ssid R.Cunning steponeR function:

* Get list of plate files to read
* Calculate the ratios
* Extract the results

```{r CalculateRatios}

  Ssid.plates <- list.files(path="Data", pattern=".csv", 
                           full.names=TRUE, recursive = TRUE)
  # Ssid.plates

  # Run stepone function to get Ratios
  
  Ssid.Out <- steponeR(files=Ssid.plates, target.ratios=c("C.Ssid", "D.Ssid"), 
                       fluor.norm=list(C=0, D=0, Ssid=0),
                       copy.number=list(C=1, D=1, Ssid=1),
                       ploidy=list(C=1, D=1, Ssid=2),
                       extract=list(C=0.813, D=0.813, Ssid=0.982))
  
  # Target ratio results
  Ssid<-Ssid.Out$result
  
```

# 2. Data CLEANING

```{r DataCleaning}
  # 1. Check and remove NTC wells
    ntc <- Ssid[which(Ssid$Sample.Name=="NTC"), ]
    Ssid <- droplevels(Ssid[!rownames(Ssid) %in% rownames(ntc), ])
    
  # 2. Check and remove + Control wells
    Positive <- Ssid[which(Ssid$Sample.Name=="+"| Ssid$Sample.Name=="Sample 1"), ]
    Ssid <- droplevels(Ssid[!rownames(Ssid) %in% rownames(Positive), ])
    
  # 3. Create unique sample ID+FileName to relabel samples  
    Ssid$Sample.Plate<-paste(Ssid$Sample.Name,Ssid$File.Name, sep="_")

  # 4.If Clade only detected in one technical replicate, set its ratio to NA
    One.C<- Ssid[which(Ssid$C.reps==1),]
    Ssid$C.Ssid[which(Ssid$C.reps==1)] <- NA
    
    One.D<- Ssid[which(Ssid$D.reps==1),]
    Ssid$D.Ssid[which(Ssid$D.reps==1)] <- NA

  # 5. Make NA=0
     # colnames(Ssid)[which(colnames(Ssid) %in% "A.Ssid")] <- "A.SH"  
     Ssid$C.Ssid[is.na(Ssid$C.Ssid)] <- 0
     Ssid$D.Ssid[is.na(Ssid$D.Ssid)] <- 0
        
  # 6.If coral detected in one technical replicate, remove the sample
    ReRun.Coral <- Ssid[which(Ssid$Ssid.reps==1), ]
    Ssid <- droplevels(Ssid[!rownames(Ssid) %in% rownames(ReRun.Coral), ])
   
```
    
# 3. Get the cell ratios and log 10 transformations

```{r}
    # Total SH
        Ssid$TotalSH<-(Ssid$C.Ssid + Ssid$D.Ssid)
    
    # Log 10
        Ssid$logC.SH <- log10(Ssid$C.Ssid)
        Ssid$logD.SH <- log10(Ssid$D.Ssid)
        Ssid$logSH<-log10(Ssid$TotalSH)
        

```    

Data summary 

```{r Summary} 

summary(Ssid)
# StepOne and Quanta mean CT values and mean SD  

Ssid %>%
  group_by(File.Name) %>%
  summarise(meanC = mean(C.CT.mean, na.rm=TRUE),
            meanD = mean(D.CT.mean, na.rm=TRUE),
            meanSsid = mean(Ssid.CT.mean),
            SD_C = mean(C.CT.sd, na.rm=TRUE), 
            SD_D = mean(D.CT.sd, na.rm=TRUE), 
            Sd_Ssid = mean(Ssid.CT.sd), n = n())

 # density Plots
        
    DenSH<- ggplot(Ssid, aes(TotalSH, fill = File.Name , colour = File.Name)) +
            geom_density(alpha = 0.1)
    DenSH
      
    DenSidCT<-ggplot(Ssid, aes(Ssid.CT.mean, fill = File.Name , colour = File.Name)) +
            geom_density(alpha = 0.1)
    DenSidCT
    
    DenC_CT<-ggplot(Ssid, aes(C.CT.mean, fill = File.Name , colour = File.Name)) +
            geom_density(alpha = 0.1)
    DenC_CT
    
    DenD_CT<-ggplot(Ssid, aes(D.CT.mean, fill = File.Name , colour = File.Name)) +
            geom_density(alpha = 0.1)
    DenD_CT

```

# 4. Tests for "teatment" = machine effects 

```{r Test}

# Total (log10) SH 
    # T test
        PLate_SH <-t.test(logSH ~ File.Name, data = Ssid)   
        PLate_SH
        
    # or (lm with Sample nested?)  
        library(nlme)
        
        SH_LM<-lme(logSH ~ File.Name, random=~1|Sample.Name,
                    data=Ssid,
                    method="REML")
        anova.lme(SH_LM)
        summary(SH_LM)
       # coef(PLate_SH)
       # layout(matrix(1:4,2,2))  
       # plot(PLate_SH)  
    
    logSH <- ggplot(Ssid, aes (File.Name, logSH, colour=factor(File.Name))) +
      stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
      stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) + theme_gdocs() +
      geom_jitter()
    logSH  

# Ssid (SYBR) CT means
    
    # T test
        PLate_Ssid <-t.test(Ssid.CT.mean ~ File.Name, data = Ssid)   
        PLate_Ssid

      Ssid_SYBER <- ggplot(Ssid, aes (File.Name, Ssid.CT.mean, colour=factor(File.Name))) +
        stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
        stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) + theme_gdocs() +
        geom_jitter()
      Ssid_SYBER


# C (VIC) CT means
      
    # T test
      PLate_C <-t.test(C.CT.mean ~ File.Name, data = Ssid) 
      PLate_C
   
      C_VIC <- ggplot(Ssid, aes (File.Name, C.CT.mean, colour=factor(File.Name))) +
        stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
        stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) + theme_gdocs() +
        geom_jitter()
      C_VIC
  
 # D (FAM) CT means
      
    # T test
    
      PLate_C <-t.test(D.CT.mean ~ File.Name, data = Ssid) 
      PLate_C
      
      D_VIC <- ggplot(Ssid, aes (File.Name, D.CT.mean, colour=factor(File.Name))) +
        stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
        stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) + theme_gdocs() +
        geom_jitter()
      D_VIC
```


## 5. Exploratory graphs - Comparision bw rxn in each sample: 

```{r Graphs}

library("ggthemes")
library("scales")

Tot_SH <- ggplot(Ssid, aes(Sample.Name , TotalSH)) +
  geom_point(aes(colour=factor(File.Name)))  + theme_gdocs() + 
  xlab("Sample") + ylab("S:H cell ratio") + theme(axis.text.x = element_blank())
Tot_SH

LogTot_SH <- ggplot(Ssid, aes(Sample.Name , logSH )) +
  geom_point(aes(colour=factor(File.Name)))  + theme_gdocs() + 
  xlab("Sample") + ylab("Log10 (S:H cell ratio)") + theme(axis.text.x = element_blank())
LogTot_SH

C_SH <- ggplot(Ssid, aes(Sample.Name , C.Ssid)) +
   geom_point(aes(colour=factor(File.Name)))  + theme_gdocs() +
   xlab("Sample") + ylab("C:H cell ratio") + theme(axis.text.x = element_blank())
C_SH

D_SH <- ggplot(Ssid, aes(Sample.Name , D.Ssid)) +
   geom_point(aes(colour=factor(File.Name)))  + theme_gdocs()+
  xlab("Sample") + ylab("D:H cell ratio") + theme(axis.text.x = element_blank())
D_SH
 
Ssid_CT <- ggplot(Ssid, aes(Sample.Name , Ssid$Ssid.CT.mean)) +
   geom_point(aes(colour=factor(File.Name)))  + theme_gdocs()  +
   xlab("Sample") + theme(axis.text.x = element_blank()) + scale_y_continuous(name="Ssid CT mean", breaks=seq(19,40,0.5))
Ssid_CT
 
C_CT <- ggplot(Ssid, aes(Sample.Name , Ssid$C.CT.mean)) +
   geom_point(aes(colour=factor(File.Name)))  + theme_gdocs()  +
   xlab("Sample") + theme(axis.text.x = element_blank()) + scale_y_continuous(name="C CT mean", breaks=seq(19,40,1))
C_CT
 
D_CT <- ggplot(Ssid, aes(Sample.Name , Ssid$D.CT.mean)) +
   geom_point(aes(colour=factor(File.Name)))  + theme_gdocs() +
   xlab("Sample")+ theme(axis.text.x = element_blank()) + scale_y_continuous(name="D CT mean", breaks=seq(19,40,1))
D_CT

```
# Differences bw the CTs in each machine

``` {r}

Data.Wide<- reshape(Ssid, idvar = "Sample.Name", timevar = "File.Name", direction = "wide")

# CTs Differences
    
    Data.Wide$Ssid_Diff<-(Data.Wide$Ssid.CT.mean.Quanta.csv-Data.Wide$Ssid.CT.mean.StepOne.csv)
      summary(Data.Wide$Ssid_Diff)
    
    Data.Wide$C_Diff<-(Data.Wide$C.CT.mean.Quanta.csv-Data.Wide$C.CT.mean.StepOne.csv)
      summary(Data.Wide$C_Diff)
      
    Data.Wide$D_Diff<-(Data.Wide$D.CT.mean.Quanta.csv-Data.Wide$D.CT.mean.StepOne.csv)
      summary(Data.Wide$D_Diff)
      
  # Graphs 
      
    Ssid_CT_Diff_A <- ggplot(Data.Wide, aes(Ssid.CT.mean.StepOne.csv , Data.Wide$Ssid_Diff)) +
          geom_point()  + theme_gdocs() + geom_hline(yintercept =0) +
          xlab("StepOne_Ssid CT") +  geom_smooth(method="lm") +
          scale_y_continuous(name=" Ssid (SYBR) CT difference (Quanta-StepOne)",breaks=seq(-2,3,0.5))
    Ssid_CT_Diff_A
    
    SYBR_CT<-lm (Ssid_Diff ~ Ssid.CT.mean.StepOne.csv, data=Data.Wide)
    summary(SYBR_CT)
    
    Ssid_CT_Diff <- ggplot(Data.Wide, aes(Ssid.CT.mean.Quanta.csv , Data.Wide$Ssid_Diff)) +
          geom_point()  + theme_gdocs() + geom_hline(yintercept =0) +
          xlab("StepOne_Ssid CT") +  geom_smooth(method="lm") +
          scale_y_continuous(name=" Ssid (SYBR) CT difference (Quanta-StepOne)",breaks=seq(-2,3,0.5))
    Ssid_CT_Diff
    
    C_CT_Diff <- ggplot(Data.Wide, aes(C.CT.mean.StepOne.csv , Data.Wide$C_Diff)) +
          geom_point()  + theme_gdocs() + geom_hline(yintercept =0) + xlab("StepOne_C CT") +  
          scale_y_continuous(name="C (VIC) CT difference (Quanta-StepOne)", breaks=seq(-2,6,0.5))
    C_CT_Diff
        
    D_CT_Diff <- ggplot(Data.Wide, aes(D.CT.mean.StepOne.csv , Data.Wide$D_Diff)) +
          geom_point()  + theme_gdocs()  + geom_hline(yintercept =0) + xlab("StepOne_D CT") + 
          scale_y_continuous(name="D (FAM) CT difference (Quanta- StepOne)", breaks=seq(-3,1,0.5))
    D_CT_Diff
        
    
# SH Differences    

# Data.Wide$C_Host_Diff<-(Data.Wide$C.Ssid.Quanta.csv-Data.Wide$C.Ssid.StepOne.csv)
# Data.Wide$D_Host_Diff<-(Data.Wide$D.Ssid.Quanta.csv-Data.Wide$D.Ssid.StepOne.csv)
# Data.Wide$TotalSH_Diff<-(Data.Wide$Ssid.CT.mean.Quanta.csv-Data.Wide$Ssid.CT.mean.StepOne.csv) 
# 
# SH_Diff <- ggplot(Data.Wide, aes(Sample.Name, TotalSH_Diff)) +
#        geom_point()  + theme_gdocs() + geom_hline(yintercept =0) +
#       xlab("Sample") + theme(axis.text.x = element_blank()) +
#       scale_y_continuous(name="SH cell ratio difference (Quanta-StepOne)",   breaks=seq(-0.5,1,0.2))
# SH_Diff
#     
# CH_Diff <- ggplot(Data.Wide, aes(Sample.Name,  C_Host_Diff)) +
#       geom_point() + xlab("Sample") +  theme(axis.text.x = element_blank()) +
#       scale_y_continuous(name="CH cell ratio difference (Quanta-StepOne)", breaks=seq(-0.5,1,0.2)) +
#       theme_gdocs() + geom_hline(yintercept =0)
# CH_Diff
    
```
    

